# Interactive component

## Load Data

The dataset contains the characteristics of the 19,000 players contained in the football simulation videogame *FIFA 22*. It can be downloaded here : [https://sports-statistics.com/soccer/fifa-2022-dataset-csvs/]

```{r}

df <- a
head(df)
```

## Cleaning
Our goal is to compare the players in a quantitative way. We will therefore remove all the qualitative data from the dataset, including the URLs. One could keep the club informations, but I decided not to because this too informative feature was hiding the real underlying patterns. Finally, I removed the goalkeepers because they were the cause of missing data.

```{r}

data_field <- data_field[data_field['player_positions']!='GK']
head(data_field)



```


```{r}

useful_cols <- 


defaultW <- getOption("warn")
options(warn = -1)


clean_col=function(dataset, col){
  column = dataset[, col]
  
  str_to_num = function(x){
    return(as.numeric(substr(x, 1, 2)))
  }
  
  return(sapply(column, FUN = str_to_num))
}


for (col in c("ls","st","rs","lw","lf","cf","rf","rw","lam","cam","ram","lm","lcm","cm","rcm","rm","lwb","ldm","cdm","rdm","rwb","lb","lcb","cb","rcb","rb","gk")){
    #print(col)
    all_players[,col]=clean_col(all_players, col)
    male_players[,col]=clean_col(male_players, col)
    female_players[,col]=clean_col(female_players, col)
    male_22[,col]=clean_col(male_22, col)
    fem_22[,col]=clean_col(fem_22, col)
}
```